HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section \ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
\newcommand{\bh}[2]{\footnote{\color{blue}BH: #1}}

@@

  Recent interest in semantics-aware MT systems underscores the importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepencies
  between the translation output and the source, is likely support the construction
  and tuning of semantics-based MT, much in the same way that string-based measures,
  such as BLEU, enabled progress in shallower MT.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating HUME's broad applicability,
  as well as its reliability, reflected in its inter-annotator agreement rates.

@@

We propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
a human evaluation measure that decomposes over the semantic units of the sentence.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

@@

We conduct experiments with four language pairs: English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
agreement scores which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's scores do not decrease with longer sentences.\bh{We may
  have to focus more on qualitative aspects, if we just try to sell
  the measure on IAA, then the figures are not stunning.}
The paper concludes by outlining ongoing work to construct semi- and
fully-automatic variants of HUME.

@@

Scalability is addressed by employing automatic and semi-automatic approximations of human
judgements. Most commonly, such scores decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a reference translation,
whose validity has been manually confirmed. This decomposition further allows system developers
to localize the errors within the translation.
The most commonly used measures decompose over n-grams or even individual words, e.g., 
the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
Another common approach is to determine the similarity between the reference and translation
by applying string edit distance \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
the evaluation of massive-scale experiments,
their decomposition over words and n-grams is better suited for localizing errors
in the output of shallow MT methods, and is less instructive in identifying the semantic
nature of the errors.

@@

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: http://www.cs.huji.ac.il/~oabend/ucca.html}.
Let $L$ be the set of possible UCCA labels.
Given a sentence $s = (w_1, \ldots,w_n)$, the UCCA annotation is a DAG $G=(V,E,\ell)$,
where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
The semantic units for a sentence $s$, induced by the annotation $G$
is the set of yields of nodes in $V$. We denote it with $Units(s)$ for brevity.
We note that the translation need not be annotated.

@@

We note that the defined labels are more similar in spirit to adequacy,
rather than fluency judgements. Specifically, annotators are instructed to
label a unit as Adequate if its translation is understandable and preserves
the meaning of the source unit, even if its fluency is impaired.

For example, consider the sentence ``After graduation, Tom moved
to NYC'' \figref{fig:ucca_example}. The atomic units are the inidividual words,
nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation
is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
  be in English for presentation purposes. In our experiments $t$ is in German,
  Czech, Polish or Romanian.}.
Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
expresses the relation between the graduation and moving Scenes, is marked Bad,
as their relation in $t$ is reversed.

>>>>

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}

@@

  Recent interest in semantics-aware MT systems underscores the importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepancies
  between the translation output and the source, is likely support the construction
  and tuning of semantics-based MT, much in the same way that string-based measures,
  such as BLEU, enabled progress in shallower MT.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT Evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating HUME's broad applicability,
  as well as its reliability, reflected in its inter-annotator agreement rates.

@@

We propose the Human UCCA-based MT Evaluation (HUME) metric,
a human evaluation measure that decomposes over the semantic units of the sentence.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

@@

We conduct experiments with four language pairs: English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. The results show that HUME obtains 
agreement scores which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's scores do not decrease with longer sentences.\bh{We may
  have to focus more on qualitative aspects, if we just try to sell
  the measure on IAA, then the figures are not stunning.}
The paper concludes by outlining ongoing work to construct semi- and
fully-automatic variants of HUME.

@@

Scalability is addressed by employing automatic and semi-automatic approximations of human
judgements. Most commonly, such scores decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a reference translation,
whose validity has been manually confirmed. This decomposition further allows system developers
to localize the errors or at least unconfirmed parts within the translation.
The most commonly used measures decompose over n-grams or even individual words, e.g., 
the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
Another common approach is to determine the similarity between the reference and translation
by applying string edit distance \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
the evaluation of massive-scale experiments,
their decomposition over words and n-grams is better suited for localizing errors
in the output of shallow MT methods, and is less instructive in identifying the semantic
nature of the errors.

@@

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
Let $L$ be the set of possible UCCA labels.
Given a sentence $s = (w_1, \ldots,w_n)$, the UCCA annotation is a DAG $G=(V,E,\ell)$,
where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
The semantic units for a sentence $s$, induced by the annotation $G$
is the set of yields of nodes in $V$. We denote it with $Units(s)$ for brevity.
We note that the translation need not be annotated.

@@

XXX: Could someone review this for Ondrej:
Note that for non-configuration languages, the relationship between the
governing and the dependent units is often expressed using morphological
properties of the dependents instead of their ordering (this holds especially for the verb and its
modifiers, less so for components of noun phrases). Since errors in morphology
should be expressed using the traffic lights on the dependent nodes, we would expect
a higher proportion of ``Adequate'' structural units for Czech and Polish.

We defined the labels to reflect more adequacy,
rather than fluency judgements. Specifically, annotators are instructed to
label a unit as Adequate if its translation is understandable and preserves
the meaning of the source unit, even if its fluency is impaired.

For example, consider the sentence ``After graduation, Tom moved
to NYC'' \figref{fig:ucca_example}. The atomic units are the individual words,
nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation
is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
  be in English for presentation purposes. In our experiments $t$ is in German,
  Czech, Polish or Romanian.}.
Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
expresses the relation between the graduation and moving Scenes, is marked Bad,
as their relation in $t$ is reversed.
